### 2013-07-10 inst/doc/reml.Rmd
updated draft outline

Consequently ecological
data reuse remains difficult and haphazard. __@mbjones: is that unfair?__

@@

[Reichman 2011]: http://dx.doi.org/10.1126/science.1197962
[Jones 2006]: http://dx.doi.org/10.1146/annurev.ecolsys.37.091305.110031
[KNB]: http://knb.ecoinformatics.org
[Berkley 2001]: http://dx.doi.org/10.1109/4236.957896
[Michener 1997]: http://dx.doi.org/10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2
[Morell 1996]: http://dx.doi.org/10.1126/science.273.5275.569
[Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3

[Michener 2006]: http://dx.doi.org/10.1016/j.ecoinf.2005.08.004
[Michener 2012]: http://dx.doi.org/10.1016/j.tree.2011.11.016

1. Commonly used data structures in R provide basic metadata automatically
2. We can more easily avoid redundant metadata entry 
3. Automate quality control checks

4. Publish
5. Search 
6. Build a synthesis platform

### Challenges

@@

The EML schema already provides some

Semantics.

>>>>

Consequently ecological
data reuse remains difficult and haphazard.

@@

[Reichman 2011]: http://dx.doi.org/10.1126/science.1197962
[Jones 2006]: http://dx.doi.org/10.1146/annurev.ecolsys.37.091305.110031
[KNB]: http://knb.ecoinformatics.org
[Berkley 2001]: http://dx.doi.org/10.1109/4236.957896
[Michener 1997]: http://dx.doi.org/10.1890/1051-0761(1997)007[0330:NMFTES]2.0.CO;2
[Morell 1996]: http://dx.doi.org/10.1126/science.273.5275.569
[Canhos 2004]: https://journals.ku.edu/index.php/jbi/article/viewArticle/3 
[Michener 2006]: http://dx.doi.org/10.1016/j.ecoinf.2005.08.004
[Michener 2012]: http://dx.doi.org/10.1016/j.tree.2011.11.016

1. Commonly used data structures in R provide basic metadata automatically
2. We can more easily avoid redundant metadata entry 
3. Automate quality control checks
4. Publish
5. Arbitrarily complex search 
6. Building a synthesis platform

### Automatic metadata detection

Not sure this counts as a section - but we can automatically detect at least the difference between numeric and character data, factors, ordered factors, etc.

### Reusing fields

- Examples of filling in contact information, etc using `eml$set`, using an existing EML file as a template, and using an external yaml file.

- Similar example for `column_metadata` and `unit_metadata` based on an existing EML file (just using `eml_read`).

### Automatic Quality Control checks

<!-- Not sure if we'll actually have/need to have all this in place for this paper?  -->
Examples of validation against schema.  Examples checking md5sums on files or other corruption checks?  Examples checking species names (taxize?)

### Publish data

Examples of commands for publishing EML files to external repositories.

### Complex search queries

Illustrate how to query for data

### A synthesis platform

Illustrate a basic example of synthesis.  While arbitrary synthesis will require semantics, a lab could store their data in EML format and build up there own library of common scripts to extract the data they need. As a lab database grows, the same queries can continue to deliver more data.  By scripting the data access, analyses can be automatically updated as the data set grows.

Consider simple example in regression between different morphological characters of species, or such.

## Challenges

@@

<!-- Example illustrating the need for semantics to resolve differences -->

### Implementation and architecture

The `reml` package is built for the Ecological Markup Language schema, [version 2.1.1], using the R software environment and the XML package for parsing and generating XML [Temple Lang 2013, XML].

<!-- Possibly cite all dependencies and suggests from the DESCRIPTION --> 
<!-- Reference any installation and configuration instructions necessary (e.g. API credentials for figshare) -->

[version 2.1.1]: http://knb.ecoinformatics.org/software/eml/eml-2.1.1
[Temple Lang 2013, XML]: http://cran.r-project.org/web/packages/XML/

### Quality control

The `reml` package can perform three levels of validation on EML files written from or read into R: validating the XML, validating against the EML 2.1.1 Schema, and performing

Several mechanisms are in place to evaluate the quality of the `reml` package functions themselves.  In addition to the examples from the documentation and the automated package checks provided for the R system, the `reml` package includes a complete test suite using the testthat package [Wickham 2011].

[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html

### Software reuse and support

The `reml` package is available through the CRAN repository under a CC0 license. At the time of writing, the current version is `r packageDescription("reml", fields="Version")`. The current development branch can be found on Github at [https://github.com/ropensci/reml].  See the NEWS file for changes in recent versions. Users are encouraged to submit bug reports or feature requests in the issues log.  Contact ropensci-discuss@googlegroups.com or the package maintainer for help.

There are many promising directions for further development of the `reml` package.  The project's Milestones page provides a list of steps currently planned or under consideration.  In particular, these milestones include increased support for interactive mode of entering data (wizards), and support for integration of semantics defined through the Web Ontology language (OWL), as discussed in the Challenges section.

### Acknowledgements

This project was supported in part by a grant from the Alfred P Sloan Foundation (CB and KR), NSF Grant DBI-1306697 (CB), ... `reml` is part of the rOpenSci project, [http://ropensci.org].

OLD OUTLINE
----------------------------------------------------------------------------------------------
